---
output:
  html_document: default
  pdf_document: default
---
# Red Wine Exploration
### by David Vartanian

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r echo=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(dplyr)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data <- read.csv("data/wineQualityReds.csv")
```

```{r echo=FALSE, plot_functions}
plot_histogram <- function(x, xlabel = '', colour = I('#6688CC'), title = '', bins = 20) {
  plot <- ggplot(aes(x = x), data = data) +
    geom_histogram(aes(colour = colour), bins = bins) +
    xlab(xlabel) +
    ggtitle(title)
  return(plot)
}

plot_outliers <- function(y, ylabel = '', title = '') {
  plot <- ggplot(aes(x = 1, y = y), data = data) +
    geom_boxplot(colour = 'blue', outlier.shape = 1, outlier.colour = 'red') +
    geom_jitter(aes(x = 1, y = y), alpha = 0.3) +
    labs(x = '', y = ylabel, 
       title = title)
  return(plot)
}

plot_quality_box <- function(y, ylabel = '', title = '') {
  plot <- ggplot(aes(x = factor(quality), y = y), data = data) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, colour = 'blue') +
  labs(x = 'Quality', y = ylabel, title = title)
  return(plot)
}
```

# Abstract
I describe a dataset with almost 1600 types of red wine, in order to understand
the meaning of the assigned score.

# Introduction
This dataset is provided by Paulo Cortez, António Cerdeira, Fernando Almeida,
Telmo Matos, and José Reis, from different universities in Portugal. It provides
information like acidity, residual sugar, chlorides, and alcohol among others.
I explore the data to find patterns and trends and get the meaning of the given
features. More information [here](http://www.sciencedirect.com/science/article/pii/S0167923609001377?via%3Dihub).


# Univariate Plots Section
Let's start showing some summary numbers and first histograms to understand
individual variables.

## Quality
This document is all about quality. Here is the distribution of wine by quality.

```{r echo=FALSE, Quality_Distribution}
ggplot(aes(x = as.factor(quality)), data = data) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 700, by = 50)) +
  labs(x = 'Quality', y = 'Count', title = 'Wine Quality Distribution')
```

### Histograms: quality, fixed.acidity, total.sulfur.dioxide, alcohol
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
# volatile.acidity
plot.volatile.acidity <- plot_histogram(data$volatile.acidity, 
                                        'Acetic Acid (g/dm³)', I('#CC9933'))
plot.volatile.acidity.outliers.box <- plot_outliers(data$volatile.acidity, 
                                                    'Acetic Acid (g/dm³)', 
                                                    'Acetic Acid Outliers') 

# fixed.acidity
plot.fixed.acidity <- plot_histogram(data$fixed.acidity, 
                                     'Tartaric Acid (g/dm³)', I('#99CCFF'))

plot.fixed.acidity.outliers.box <- plot_outliers(data$fixed.acidity,
                                                 'Tartaric Acid (g/dm³)',
                                                 'Tartaric Acid Outliers')

# Sulphure Dioxide
plot.total.sulfur.dioxide <- plot_histogram(data$total.sulfur.dioxide,
                                            'Total Sulphure Dioxide (mg/dm³)',
                                            I('#00DD99'))
plot.total.sulfur.dioxide.outliers.box <- plot_outliers(
  data$total.sulfur.dioxide,
  'Total Sulphure Dioxide (mg/dm³)',
  'Total Sulphure Dioxide Outliers')

# Alcohol
plot.alcohol <- plot_histogram(data$alcohol, 'Alcohol (%)',
                               I('#FF3366'), 10)
plot.alcohol.outliers.box <- plot_outliers(
  data$alcohol, 'Alcohol (%)', 'Alcohol Outliers')

# Chlorides
plot.chlorides <- plot_histogram(data$chlorides, 'Sodium Chloride (g/dm³)',
                                 I('#CCFF00'), bins = 10)
plot.chlorides.outliers.box <- plot_outliers(
  data$chlorides,
  'Sodium Chloride (g/dm³)', 'Sodium Chloride Outliers')

# Residual Sugar
plot.residual.sugar <- plot_histogram(data$residual.sugar, 
                                      'Residual Sugar (g/dm³)', I('#6688CC'))
plot.residual.sugar.outliers.box <- plot_outliers(
  data$residual.sugar, 'Residual Sugar (g/dm³)', 'Residual Sugar Outliers')

# pH
plot.pH <- plot_histogram(data$pH, 'pH', I('#CC00CC'))
plot.pH.outliers.box <- plot_outliers(data$pH, 'pH', 'pH Outliers')

# Sulphates
plot.sulphates <- plot_histogram(data$sulphates, 
                                 'Potassium Sulphate (g/dm³)', 
                                 I('#33DDDD'), bins = 10)
plot.sulphates.outliers.box <- plot_outliers(
  data$sulphates,
  'Potassium Sulphate (g/dm³)',
  'Potassium Sulphate Outliers')

# Density
plot.density <- plot_histogram(data$density, 'Density of water', I('#91C0F9'))
plot.density.outliers.box <- plot_outliers(data$density, 
                                           'Density of water',
                                           'Density Outliers')

grid.arrange(plot.volatile.acidity, plot.fixed.acidity, 
             plot.total.sulfur.dioxide, plot.alcohol, 
             plot.chlorides, plot.residual.sugar, plot.pH, plot.sulphates, 
             plot.density, ncol = 3)
```

These histograms show how the values are distributed in the different variables.

### Outliers & Statistical Info

```{r echo=FALSE, warning=FALSE, error=FALSE, Outliers1}
grid.arrange(plot.volatile.acidity.outliers.box, 
             plot.volatile.acidity, 
             ncol = 2)
```

```{r}
# Statistical information about Volatile Acidity
summary(data$volatile.acidity)
```

There are a few outliers only on the right side.

```{r echo=FALSE, warning=FALSE, error=FALSE, Outliers2}
grid.arrange(plot.fixed.acidity.outliers.box, plot.fixed.acidity, ncol = 2)
```

```{r}
# Statistical information about Fixed Acidity
summary(data$fixed.acidity)
```

There are several outliers only on the right side.

```{r echo=FALSE, Outliers3}
grid.arrange(plot.total.sulfur.dioxide.outliers.box, 
             plot.total.sulfur.dioxide, 
             ncol = 2)
```

```{r}
# Statistical information about Total Sulphure Dioxide
summary(data$total.sulfur.dioxide)
```

There are many outliers only on the right side.

```{r echo=FALSE, Outliers4}
grid.arrange(plot.alcohol.outliers.box, plot.alcohol, ncol = 2)
```

```{r}
# Statistical information about Alcohol
summary(data$alcohol)
```

There are just a few outliers only on the right side.

```{r echo=FALSE, Outliers5}
grid.arrange(plot.chlorides.outliers.box, plot.chlorides, ncol = 2)
```

```{r}
# Statistical information about Chlorides
summary(data$chlorides)
```

There are just a few outliers on the left side, and many on the right side.

```{r echo=FALSE, Outliers6}
grid.arrange(plot.residual.sugar.outliers.box, plot.residual.sugar, ncol = 2)
```

```{r}
# Statistical information about Residual Sugar
summary(data$residual.sugar)
```

There are many outliers on the right side.

```{r echo=FALSE, Outliers7}
grid.arrange(plot.pH.outliers.box, plot.pH, ncol = 2)
```

```{r}
# Statistical information about pH
summary(data$pH)
```

All values are pretty well distributed in the *pH* variable. There are several 
outliers on both sides.

```{r echo=FALSE, Outliers8}
grid.arrange(plot.sulphates.outliers.box, plot.sulphates, ncol = 2)
```

```{r}
# Statistical information about Sulphates
summary(data$sulphates)
```

There are many outliers only on the right side.

```{r echo=FALSE, Outliers9}
grid.arrange(plot.density.outliers.box, plot.density, ncol = 2)
```

```{r}
# Statistical information about Density
summary(data$density)
```

This variable is also well distributed. There are several outliers on both 
sides.

# Univariate Analysis

### Dataset Structure
There are 9 continuous variables, 2 discrete variables and one ordered 
categorical variable: *quality*.

### Main dataset interest
My general question is, how do chemical properties define the quality of the red 
wine?

There are interesting features in this dataset, each of them describing an
important property of the red wine. Density, pH, sulphur dioxide, and sulphates
are, in my opinion, the most important ones, in order to measure the quality.
Let's see what we can find by looking at those variables.

### Variable Transformations
It was not necessary to clean missing values on this dataset. However, I think 
it is a good idea to apply some transformations to skewed variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables1}
# create categorical variable: durability
data$durability <- ifelse(data$sulphates > median(data$sulphates) 
                          & data$free.sulfur.dioxide > median(
                            data$free.sulfur.dioxide), 
                          c("Long"), c("Short"))
grid.arrange(plot.volatile.acidity + 
               xlab('Acetic Acid - No transformation'), 
             plot.volatile.acidity + xlab('Acetic Acid - Log10') +
               scale_x_log10(breaks = seq(0,1.6,0.4)), ncol = 2, 
             top = 'Acetic Acid Tranformation')
```


Transformed Volatile Acidity using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables2}
grid.arrange(plot.fixed.acidity + xlab('Tartaric Acid - No transformation'), 
             plot.fixed.acidity + xlab('Tartaric Acid - Log10') +
               scale_x_log10(breaks=seq(4,15,1)), ncol = 2, 
             top = 'Tartaric Acid Tranformation')
```

Transformed Tartaric Acid using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables3}
grid.arrange(plot.total.sulfur.dioxide + 
               xlab('Total Sulphure Dioxide - No transformation'), 
             plot.total.sulfur.dioxide + 
               xlab('Total Sulphure Dioxide - Log10') +
               scale_x_log10(breaks = seq(8, 320, 80)), ncol = 2, 
             top = 'Total Sulphure Dioxide Tranformation')
```

Transformed Total Sulphure Dioxide using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables4}
grid.arrange(plot.chlorides + xlab('Sodium Chloride - No transformation'), 
             plot.chlorides + xlab('Sodium Chloride - Log10') +
               scale_x_log10(breaks = seq(0,0.6,0.2)), ncol = 2, 
             top = 'Sodium Chloride Tranformation')
```

Transformed Chlorides using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables5}
grid.arrange(plot.residual.sugar + xlab('Residual Sugar - No transformation'), 
             plot.residual.sugar + xlab('Residual Sugar - Log10') +
               scale_x_log10(breaks = seq(0, 14, 2)), ncol = 2, 
             top = 'Residual Sugar Tranformation')
```

Transformed Residual Sugar using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables6}
grid.arrange(plot.sulphates + xlab('Potassium Sulphate - No transformation'), 
             plot.sulphates + xlab('Potassium Sulphate - Log10') +
               scale_x_log10(breaks = seq(0, 2, 0.4)), ncol = 2, 
             top = 'Potassium Sulphate Tranformation')
```

Transformed Sulphates using log base 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables7}
plot.free.sulfur.dioxide <- ggplot(aes(x = free.sulfur.dioxide), data = data) +
  geom_histogram(aes(colour = I('#00DD99'))) +
  xlab('Free Sulphure Dioxide (mg/dm³)')
grid.arrange(plot.free.sulfur.dioxide + 
               xlab('Free Sulphure Dioxide - No transformation'), 
             plot.free.sulfur.dioxide + 
               xlab('Free Sulphure Dioxide - Log10') +
               scale_x_log10(breaks = c(1,10,20,35,50,70)), ncol = 2, 
             top = 'Free Sulphure Dioxide Tranformation')
```

Transformed Free Sulphure Dioxide using log base 10.

#### Durability
Using the new variable *durability*, it's possible to appreciate the effect 
of *sulphates* and *free sulphure dioxide*.

This variable has only two values: S (short) and L (large), using the median of
*sulphates* and *free sulphure dioxide* as inflection point.

# Bivariate Plots Section
Let's try to find trends and interesting patterns by comparing two variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Distribution_Alcohol}
plot_quality_box(data$alcohol, 'Alcohol (%)', 'Alcohol level by Quality')
```

```{r}
cor(x = data$quality, y = data$alcohol)
```

> **Fact**: Higher quality wines seem to have higher levels of alcohol

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Dist_FixedAcidity}
plot_quality_box(data$fixed.acidity, 'Tartaric Acid (g/dm³)', 
                 'Tartaric Acid level by Quality')
```

```{r}
# Correlation between Quality and Fixed Acidity
cor(x = data$quality, y = data$fixed.acidity)
```

> **Fact**: Higher quality wines seem to have lower levels of acidity

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_vs_Density}
plot_quality_box(data$density, 'Density (g/cm³)', 'Density level by Quality')
```

```{r}
# Correlation between Quality and Density
cor(x = data$quality, y = data$density)
```

> **Fact**: Higher quality wines seem to have lower density

```{r echo=FALSE, Bivariate_Plots1}
plot.citric_acid_pH <- ggplot(aes(x = citric.acid, y = pH), data = data) +
  geom_point(colour = I('#8DB7A4'), alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1, colour = 'brown')  +
  xlab('Citric Acid (g/dm³)') +
  ggtitle('Citric Acid vs. pH')
plot.citric_acid_pH
```

Citric Acid adds freshness flavor to the wine.

```{r}
# Correlation between Citric Acid and pH
cor(x = data$citric.acid, y = data$pH)
```

```{r echo=FALSE, Bivariate_Plots2}
ggplot(aes(x = volatile.acidity, y = pH), data = data) +
  geom_point(colour = I('#B19784'), alpha = 0.5) +
  xlab('Acetic Acid (g/dm³)') +
  ggtitle('Acetic Acid vs. pH')
```

Level of acetic acid. Too high levels make an unpleasant vinegar taste.

```{r}
# Correlation between Volatile Acidity and pH
cor(x = data$volatile.acidity, y = data$pH)
```

# Bivariate Analysis

### Relationships

```{r echo=FALSE, Density_vs_Quality}
ggplot(aes(x = density, y = quality, colour = I('#3A7C9D')), data = data) +
  geom_point(position = 'jitter', alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1, colour = 'brown')  +
  labs(x = 'Density (g/cm³)', y = 'Quality', title = 'Density vs. Quality')
```

```{r}
# Correlation between Density and Quality
cor(x = data$density, y = data$quality)

```

I've found a slightly positive correlation, meaning that *density* tends to be
lower on high-quality wines. However, this correlation is not so important to 
determine the quality as shown by the correlation coefficient.

```{r echo=FALSE, warnings=FALSE, Chlorides_vs_Sulphates}
ggplot(aes(x = chlorides, y = sulphates, colour = sulphates), data = data) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE,size=1, colour = 'brown')  +
  xlim(c(0,0.2)) +
  ylim(c(0, 1.5)) +
  labs(x = 'Sodium Chloride (g/dm³)', y = 'Potassium Sulphate (g/dm³)', 
       title = 'Sodium Chloride vs. Potassium Sulphate')
```

```{r}
# Correlation between Chlorides and Sulphates
cor(x = data$chlorides, y = data$sulphates)
```

I've found that levels are mostly low for both variables. I would say that they 
don't influence much on the quality as all types of wine have the same level of 
these two variables.

```{r echo=FALSE, warnings=FALSE, Residual_Sugar_vs_Chlorides}
ggplot(aes(x = chlorides, y = residual.sugar, colour = residual.sugar), 
       data = data) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE,size=1, colour = 'brown')  +
  xlim(c(0, 0.2)) +
  ylim(0, 10) +
  labs(x = 'Sodium Chloride (g/dm³)', y = 'Residual Sugar (g/dm³)', 
       title = 'Sodium Chloride vs. Residual Sugar')
```

```{r}
# Correlation between Chlorides and Residual Sugar
cor(x = data$chlorides, y = data$residual.sugar)
```

I've found the same here, as they keep levels constantly low and correlation
coefficient is almost 0.

```{r echo=FALSE, warning=FALSE, Free_Sulph_Dioxide_vs_Total_Sulphure_Dioxide}
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = data) +
  geom_point(colour = I('#2A67C4'), alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE,size=1, colour = 'brown')  +
  xlim(c(0, 200)) +
  ylim(c(0, 60)) +
  labs(x = 'Total Sulphure Dioxide (mg/dm³)', 
       y = 'Free Sulphure Dioxide (mg/dm³)', 
       title = 'Free vs. Total Sulphure Dioxide')

score.sulfur.dioxide <- (log10(data$free.sulfur.dioxide) + 
                           log10(data$total.sulfur.dioxide)) / 2
```

```{r}
# Correlation between Free Sulphure Dioxide and Total Sulphure Dioxide
cor(x = data$free.sulfur.dioxide, y = data$total.sulfur.dioxide)
```

Levels are always low. However, these two variables seem to be correlated.

### Interesting relationships
So far I find only *density* to be an interesting variable to look at.
The rest, *chlorides*, *sulphates*, *residual sugar* and *sulphure dioxide* 
don't seem to be a great influence on wine quality.

#### pH
This variable indicates the acidity level of the wine.
The scale goes from 0 (very acid) to 14 (very basic). But most of red wines are
statistically between 3 and 4.

```{r echo=FALSE, pH1}
ggplot(aes(x = pH), data = data) +
  facet_wrap(~quality) +
  geom_histogram(binwidth = 0.01, aes(fill = as.factor(quality))) +
  scale_fill_brewer(type='seq') +
  labs(x = 'pH', title = 'pH faceted by Quality')
```

Another point of view:

```{r echo=FALSE, pH2}
ggplot(aes(x = factor(quality), y = pH), data = data) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  labs(x = 'Quality', y = 'pH', title = 'pH level by Quality')
```

It's quite surprising that levels of pH are lower on high-quality wines.

#### Density of water
The level of this variable depends on alcohol percentage and sugar.

```{r echo=FALSE, Density_vs_Quality_Faceted}
plot.density.quality.faceted <- ggplot(aes(x = density), data = data) +
  facet_wrap(~quality) +
  geom_histogram(binwidth = 0.0001, aes(fill = as.factor(quality))) +
  scale_fill_brewer(type='seq') +
  labs(x = 'Density (g/cm³)', title = 'Density faceted by Quality')
plot.density.quality.faceted
```

Another point of view:

```{r echo=FALSE, Density2}
plot.density.quality.box <- ggplot(aes(x = factor(quality), y = density), 
                                   data = data) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  labs(x = 'Quality', y = 'Density (g/cm³)', title = 'Density level by Quality')
plot.density.quality.box
```

Density levels are also lower on high-quality wines.

#### Free Sulphure Dioxide
The free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved 
gas) and *bisulfite ion*. It prevents microbial growth and the oxidation of the 
wine.

```{r echo=FALSE, Free_Sulphur_dioxide}
ggplot(aes(x = free.sulfur.dioxide), data = data) +
  facet_wrap(~quality) +
  geom_histogram(binwidth = 3, aes(fill = as.factor(quality))) +
  scale_fill_brewer(type='seq') +
  labs(x = 'Free Sulphure Dioxide (mg/dm³)', 
       title = 'Free Sulphure Dioxide faceted by Quality')
```

Another point of view:

```{r echo=FALSE, Free_Sulphur_dioxide2}
ggplot(aes(x = factor(quality), y = free.sulfur.dioxide), data = data) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  labs(x = 'Quality', y = 'Free Sulphure Dioxide (mg/dm³)', 
       title = 'Free Sulphure Dioxide level by Quality')
```

Again, the levels for this variable are lower for both low-quality and 
high-quality wines.

#### Sulphates
Additive contributing with sulphure dioxide gas (S02) levels, acting as an
antimicrobial and antioxidant.

```{r echo=FALSE, Sulphur_dioxide}
ggplot(aes(x = sulphates), data = data) +
  facet_wrap(~quality) +
  geom_histogram(binwidth = 0.1, aes(fill = as.factor(quality))) +
  scale_fill_brewer(type='seq') +
  labs(x = 'Potassium Sulphate (g/dm³)', 
       title = 'Potassium Sulphate faceted by Quality')
```

```{r echo=FALSE, Sulphur_dioxide2}
ggplot(aes(x = factor(quality), y = sulphates), data = data) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  labs(x = 'Quality', y = 'Potassium Sulphate (g/dm³)', 
       title = 'Potassium Sulphate level by Quality')
```

Sulphates levels are lower for low-quality and high-quality wines as well.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots1}
plot.durability.quality_alcohol <- ggplot(aes(x = factor(quality),
           y = alcohol ), 
           data = data) + 
  labs(x = 'Quality', y = 'Alcohol (%)', 
       title = 'Durability regarding Alcohol vs. Quality') +
   geom_boxplot( aes(fill= durability))  +
  scale_fill_brewer(type='seq',
                   guide=guide_legend(title='Durability'))
plot.durability.quality_alcohol
```

A quite strong correlation can be observed between these two variables, 
regarding the quality of wines. Meaning that it's normal to find lower levels 
of pH and density on high-quality wines. The lines colours let you see how
durable the wine can be respect to *alcohol*, using the *durability* variable
introduced above. It makes sense to me that wines last longer if they contains
more alcohol in addition to *sulphates* and *free sulphure dioxide*.

```{r echo=FALSE, Multivariate_Plots2}
ggplot(aes(x=density, y=alcohol),
       data=data)+
  geom_smooth(aes(color=data$durability), method = 'loess') +
  labs(x = 'Density (g/cm³)', y = 'Alcohol (%)', 
       title = 'Durability regarding Density vs. Alcohol')
```

I've found here another interesting correlation, which becomes quite obvious if 
we pay special attention to the meaning of the variables. Density, as I said 
above, is actually density of water. So, the more alcohol the less water.
The coloured lines show that the *durability* of the wine is lower when the
*density of water* is higher. Does it make sense?

# Multivariate Analysis


```{r echo=FALSE, Summarise_By_Quality}
winesByQuality <- data %>%
  group_by(quality) %>%
  summarise(mean_quality = mean(quality),
            mean_alcohol = mean(alcohol),
            mean_density = mean(density),
            mean_ph = mean(pH),
            mean_citric_acid = mean(citric.acid),
            n = n())
summary(winesByQuality)
```

------

# Final Plots and Summary

### Durability & Alcohol
```{r echo=FALSE, Plot_One}
plot.durability.quality_alcohol
```

Something very remarkable to keep in mind is what this plot shows: high-quality
wines seem to last longer. But the orange line on the top-right corner makes
a huge difference. They last much longer when alcohol level is higher.

### Citric Acid vs. pH
```{r echo=FALSE, Plot_Two}
plot.citric_acid_pH
```

This is a pretty straight forward correlation. When *pH* level gets lower (which
means that there is more acid) *citric acid* gets higher. It makes sense, 
doesn't it?

### Density by Quality Level
```{r echo=FALSE, Plot_Three}
plot.density.quality.box
```

I wanted to emphasize this plot again because levels of density look similar for
both low-quality and high-quality wines. Or from another perspective, the 
density of water is higher only on mid-quality wines.

------

# Reflection
At the beginning it was for me quite difficult how to detect which levels of 
acids and minerals would be beneficial for the quality of wine.
Understanding univariate visualisations helped me to find relationships and go 
deeper with the exploration comparing variables and finding correlations.
Finally, after visualising bivariate plots, I could create even more advanced
visualisations being able to display the trend of one variable regarding other
two variables in the multivariate section.

I feel that now I have a few extra tips to select new wines to taste. Higher 
levels of alcohol and acidity, lower levels of density, as well as low levels of 
residual sugar, chlorides, and sulphates. High levels of alcohol and low level 
of density were definitely surprising for me.
However, I think that the data set needs some more categorical variables and 
much more data to make better analysis.

For instance, adding columns with usual customers, *sommeliers* preferences,
country of origin, types of grape, altitude of grape crops, and type of
cask used to keep them before selling would be of great value to measure wine 
quality beyond the product itself, but also the background environment and 
production process.

