Red Wine Exploration by DAVID VARTANIAN
=======================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data <- read.csv("data/wineQualityReds.csv")
```

# Abstract
I describe a dataset with almost 1600 types of red wine, in order to understand
the meaning of the assigned score.

# Introduction
This dataset provided by Paulo Cortez, António Cerdeira, Fernando Almeida,
Telmo Matos, and José Reis, from different universities in Portugal. It provides
information like acidity, residual sugat, chlorides, and alcohol among others.
I explore the data to find patterns and trends, and get the meaning of the given
fetures.


# Univariate Plots Section
Let's start showing some summary numbers and first histograms to understand
individual variables.

```{r echo=FALSE, Summary}
summary(data)
```

## Histograms: quality, fixed.acidity, total.sulfur.dioxide, alcohol
```{r echo=FALSE, Univariate_Plots}
p1 <- ggplot(aes(x=volatile.acidity), data=data)+
  geom_histogram()+
  xlab('Volatile Acidity')

p2 <- ggplot(aes(x=fixed.acidity), data=data)+
  geom_histogram(binwidth = 1)+
  xlab('Fixed Acidity')

p3 <- ggplot(aes(x=total.sulfur.dioxide), data=data)+
  geom_histogram(binwidth = 10)+
  xlab('Total Sulfur Dioxide')

p4 <- ggplot(aes(x=alcohol), data=data)+
  geom_histogram(binwidth = 0.4)+
  xlab('Alcohol %')

p5 <- ggplot(aes(x=chlorides), data=data)+
  geom_histogram()+
  xlab('Chlorides')

p6 <- ggplot(aes(x=residual.sugar), data=data)+
  geom_histogram()+
  xlab('Residual Sugar')

p7 <- ggplot(aes(x=pH), data=data)+
  geom_histogram()+
  xlab('pH')

p8 <- ggplot(aes(x=sulphates), data=data)+
  geom_histogram()+
  xlab('Sulphates')

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol=2)
```

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### Dataset Structure
There are 9 continuous variables, 2 discrete variables and one categorical
variable: quality.

### Main dataset interest
My general question is, how chemical properties define the quality of red wine?

### Other features of interest
I think fixed acidity, total sulfur dioxide, alcohol and sulphates are important
to analyse together.

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!