# Red Wine Exploration
## by David Vartanian

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data <- read.csv("data/wineQualityReds.csv")
```

# Abstract
I describe a dataset with almost 1600 types of red wine, in order to understand
the meaning of the assigned score.

# Introduction
This dataset is provided by Paulo Cortez, António Cerdeira, Fernando Almeida,
Telmo Matos, and José Reis, from different universities in Portugal. It provides
information like acidity, residual sugat, chlorides, and alcohol among others.
I explore the data to find patterns and trends, and get the meaning of the given
fetures. More information [here](http://www.sciencedirect.com/science/article/pii/S0167923609001377?via%3Dihub)


# Univariate Plots Section
Let's start showing some summary numbers and first histograms to understand
individual variables.

```{r echo=FALSE, Summary}
summary(data)
```

## Histograms: quality, fixed.acidity, total.sulfur.dioxide, alcohol
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
p1 <- ggplot(aes(x = volatile.acidity), data = data)+
  geom_histogram(aes(colour = I('#CC9933')))+
  xlab('Volatile Acidity')

p2 <- ggplot(aes(x = fixed.acidity), data = data) +
  geom_histogram(aes(colour = I('#99CCFF'))) +
  xlab('Fixed Acidity')

p3 <- ggplot(aes(x = total.sulfur.dioxide), data = data) +
  geom_histogram(aes(colour = I('#00DD99'))) +
  xlab('Total Sulfur Dioxide')

p4 <- ggplot(aes(x = alcohol), data = data) +
  geom_histogram(aes(colour = I('#FF3366'))) +
  xlab('Alcohol %')

p5 <- ggplot(aes(x = chlorides), data = data) +
  geom_histogram(aes(colour = I('#CCFF00'))) +
  xlab('Chlorides')

p6 <- ggplot(aes(x = residual.sugar), data = data) +
  geom_histogram(aes(colour = I('#6688CC'))) +
  xlab('Residual Sugar')

p7 <- ggplot(aes(x = pH), data = data) +
  geom_histogram(aes(colour = I('#CC00CC'))) +
  xlab('pH')

p8 <- ggplot(aes(x = sulphates), data = data) +
  geom_histogram(aes(colour = I('#33DDDD'))) +
  xlab('Sulphates')

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Distribution_Alcohol}
ggplot(aes(x = alcohol, y = quality), data = data) +
  geom_smooth(aes(colour = quality), method = 'loess') +
  ggtitle('Quality vs. Level of alcohol') +
  xlab('Alcohol %') +
  ylab('Quality')
```

> **Fact**: Higher quality wines seem to have higher levels of alcohol

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Distribution_FixedAcidity}
ggplot(aes(x = fixed.acidity, y = quality), data = data) +
  geom_smooth(aes(colour = quality), method = 'loess') +
  ggtitle('Quality vs. Level of fixed acidity') +
  xlab('Fixed Acidity') +
  ylab('Quality')
```

> **Fact**: Higher quality wines seem to have lower levels of acidity

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_vs_Density}
ggplot(aes(x = density, y = quality), data = data) +
  geom_smooth(aes(colour = quality), method = 'loess') +
  ggtitle('Density vs. Quality') +
  xlab('Density') +
  ylab('Quality')
```



> **Fact**: Higher quality wines seem to have lower density

# Univariate Analysis

### Dataset Structure
There are 9 continuous variables, 2 discrete variables and one categorical
variable: quality.

### Main dataset interest
My general question is, how chemical properties define the quality of red wine?

### Other features of interest
I think fixed acidity, total sulfur dioxide, alcohol and sulphates are important
to analyse together.

### Acidity Score
Since acidity has 4 variables, I've created a variable called *score.acidity*, in
order to measure how acid a wine can be.

```{r Score_Acidity}
fixed.acidity.norm <- (data$fixed.acidity - min(data$fixed.acidity)) / (max(data$fixed.acidity) - min(data$fixed.acidity))
volatile.acidity.norm <- (data$volatile.acidity - min(data$volatile.acidity)) / (max(data$volatile.acidity) - min(data$volatile.acidity))
citric.acid.norm <- (data$citric.acid - min(data$citric.acid)) / (max(data$citric.acid) - min(data$citric.acid))
pH.norm <- 1 - (data$pH - min(data$pH)) / (max(data$pH) - min(data$pH))
data$score.acidity <- (fixed.acidity.norm + volatile.acidity.norm + citric.acid.norm + pH.norm) / 4

ggplot(aes(x = score.acidity, y = quality), data = data) +
  geom_point(aes(colour = quality, alpha = 1/20), position = 'jitter') +
  ylab('Quality') +
  xlab('Acidity Score') +
  ggtitle('Acidity Score vs. Quality')

ggplot(aes(x = score.acidity, y = density), data = data) +
  geom_point(aes(colour = score.acidity, alpha = 1/10)) +
  labs(x = 'Acidity Score', y = 'Density', title = 'Acidity vs. Density')

# acidity by quality
ggplot(aes(x = score.acidity), data = data) +
  facet_wrap(~quality) +
  geom_histogram(binwidth = 0.05) +
  labs(x = 'Acidity Score', title = 'Acidity Score faceted by Quality')
```

### Variable Transformations
It was not necessary to clean the dataset. However I think it is a good idea to
apply some transformations to skewed variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, Transforming_Variables}
p1 + scale_x_log10() + ggtitle('Volatile Acidity - Log10')
summary(log10(data$volatile.acidity))
p2 + scale_x_log10() + ggtitle('Fixed Acidity - Log10')
summary(log10(data$fixed.acidity))
p3 + scale_x_log10() + ggtitle('Total Sulfur Dioxide - Log10')
summary(log10(data$total.sulfur.dioxide))
p4 + scale_x_log10() + ggtitle('Alcohol - Log10')
summary(log10(data$alcohol))
p5 + scale_x_log10() + ggtitle('Chloridres - Log10')
summary(log10(data$chlorides))
p6 + scale_x_log10() + ggtitle('Residual Sugar - Log10')
summary(log10(data$residual.sugar))
p8 + scale_x_log10() + ggtitle('Sulphates - Log10')
summary(log10(data$sulphates))
ggplot(aes(x = free.sulfur.dioxide), data = data) +
  geom_histogram(aes(colour = I('#00DD99'))) +
  xlab('Free Sulfur Dioxide') +
  ggtitle('Free Sulfur Dioxide - Log10') +
  scale_x_log10()
```

# Bivariate Plots Section
Let's try to find trends and interesting patterns by comparing two variables.

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = citric.acid, y = pH), data = data) +
  geom_point(aes(colour = I('#8DB7A4'), alpha = 1/10)) +
  xlab('Citric Acid') +
  ggtitle('Citric Acid vs. pH')

ggplot(aes(x = volatile.acidity, y = pH), data = data) +
  geom_point(aes(colour = I('#B19784'), alpha = 1/10)) +
  xlab('Volatile Acidity') +
  ggtitle('Volatile Acidity vs. pH')

ggplot(aes(x = density, y = quality), data = data) +
  geom_point(aes(colour = quality, alpha = 1/10), position = 'jitter') +
  xlab('Density') +
  ylab('Quality') +
  ggtitle('Density vs. Quality')
```

# Bivariate Analysis

### Relationships

```{r echo=FALSE, Density_vs_Quality}
ggplot(aes(x = density, y = quality), data = data) +
  geom_point(aes(colour = density, alpha = 1/10), position = 'jitter') +
  labs(x = 'Density', y = 'Quality', title = 'Density vs. Quality')
```

```{r echo=FALSE, Chlorides_vs_Sulphates}
ggplot(aes(x = chlorides, y = sulphates), data = data) +
  geom_point(aes(colour = sulphates, alpha = 1/20)) +
  labs(x = 'Chlorides', y = 'Sulphates', title = 'Chlorides vs. Sulphates')
```

```{r echo=FALSE, Residual_Sugar_vs_Chlorides}
ggplot(aes(x = chlorides, y = residual.sugar), data = data) +
  geom_point(aes(colour = residual.sugar, alpha = 1/10)) +
  labs(x = 'Chlorides', y = 'Residual Sugar', title = 'Chlorides vs. Residual Sugar')
```

```{r echo=FALSE, Free_Sulfur_Dioxide_vs_Total_Sulfur_Dioxide}
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = data) +
  geom_point(aes(colour = free.sulfur.dioxide, alpha = 1/10)) +
  labs(x = 'Total Sulfur Dioxide', y = 'Free Sulfur Dioxide', title = 'Free vs. Total Sulfur Dioxide')

score.sulfur.dioxide <- (log10(data$free.sulfur.dioxide) + log10(data$total.sulfur.dioxide)) / 2
```

### Interesting relationships

#### Density vs Quality
I've found a slightly positive correlation, meaning that *density* is not so
important to determine the quality.

#### Chlorides vs Sulphates
I've found that levels are mostly low for both variables. I would say that they
don't influence much on the quality, since all types of wine have the same level
on these two variables.

#### Chlorides vs Residual Sugar
I've found the same here, they keep levels always more or less low.

### Acidity vs. Density
Here the positive correlation is pretty strong and easily noticeable. 
This means that the more the level of density the more acid the wine is. And
regarding the level of acidity vs. quality, you might notice that the plots for
Acidity vs. Quality and Acidity vs. Density are almost identical.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
# acidity vs. density > quality
ggplot(aes(x = score.acidity, y = density), data = data) +
  geom_point(aes(colour = data$quality, alpha = 1/10))

# sulfur dioxide vs acidity > quality
ggplot(aes(x = score.sulfur.dioxide, y = score.acidity), data = data)+
  geom_jitter(aes(color = quality))

ggplot(aes(x = score.acidity, y = score.sulfur.dioxide), data = data) +
  geom_jitter(aes(colour = density)) +
  facet_wrap(~quality) +
  labs(x = 'Acidity Score', y = 'Sulfur Dioxide Score', title = 'Density on Acidity vs. Sulfur Dioxide, faceted by Quality')
```

# Multivariate Analysis
@todo

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. colour, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!